---
title: "Class 11"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Importing our PDB statistics CSV file and calculate percent strucutres by experimental method

```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)
```

Ho many total structures?

```{r}
percent <- (p$Total/ sum(p$Total)) * 100

names(percent) <- row.names(p)

percent
```

## Using bio3d

Loading bio3d package

```{r}
library(bio3d)
```

Read in our HIV-Protease structure

```{r}
pdb <- read.pdb("1HSG")

pdb

```

Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein
residues?

198 based on "residues/calpha atoms#"

Print a subset of $atom data for the first two atoms
```{r}

pdb$atom[1:2, c('eleno', 'elety', 'x', 'y', 'z')]

```

Note that individual $atom records can also be accessed like this


```{r}
pdb$atom$elety[1:2]
```

Which allows us to do the following

```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab= "B-factor")
```

What type of R object is pdb$atom? HINT: You can always use the str() function to get a
useful summery of any R object.

```{r}
class(pdb$atom)
```

Note that the main xyz coordinate attribute is a numeric matrix with 3N columns (each atom has
three values x, y and z). The number of rows here correspond to the number of models in the
PDB file (typically one for X-ray structures and multiple for NMR structures).

Print a summary of the coordinate data in $xyz

```{r}
print(pdb$xyz)

# dimensions 
dim(pdb$xyz)

# Print coordinates for the first two atom

pdb$xyz[1, atom2xyz(1:2)]

```

The Bio3D atom.select() function is arguably one of the most challenging for newcomers to
master. It is however central to PDB structure manipulation and analysis. At its most basic, this
function operates on PDB structure objects (as created by read.pdb()) and returns the
numeric indices of a selected atom subset. These indices can then be used to access the
$atom and $xyz attributes of PDB structure related objects.

**Q8** Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in
VMD. Also write out a second separate PDB file for the ligand with residue name MK1

HINT: In Bio3D you can use the trim.pdb() function together with the pub object and your atom
selections such as atom.select(pdb, ’protein’) for the fist file and atom.select(pdb, ‘ligand’) for
the second.
Page


```{r}
inds.ligand <- atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.protein
```

Note how xyz is 3 times the length of atom indices, makes sense as this give a 3d coordinate graph


```{r}
pdb$atom[inds.ligand$atom,]

```

```{r}
pdb$atom[inds.protein$atom,]
```

This is desired data, but we need to put it into a file

```{r}
pdb.ligand <- trim.pdb(pdb, inds = inds.ligand)
pdb.ligand
```

```{r}
write.pdb(pdb.ligand,file = "1hsg_ligand.pdb")
```

Same code done for protease

```{r}
pdb.protein <- trim.pdb(pdb, inds = inds.protein)
pdb.protein

write.pdb(pdb.protein, file = "1hsg_protein.pdb")
```

# Section 4 

**Testing muscle.exe on examples**
Downloading example PDB files

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

The get.pdb() function will download the requested files. Argument split = TRUE requests
further that we want to extract particular chains, i.e. those specified by the _A suffix of each
PDB ID in the example above. Note that these ids could come from the results of a
blast.pdb() search as described in the next section.

The requested chains are then aligned and their structural data stored in a new object pdbs that
can be used for further analysis. The pdbaln() function includes the ability to superimpose, or
fit, all structures onto each other using the argument fit = TRUE.

```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)

```

**Q9**: What effect does setting the fit=TRUE option have in the related rmsd() function?
What does RMSD measure and what would the results indicate if you set fit=FALSE or
removed this option? 

**Answer**: fit = TRUE performs coordinate superposition before performing the RMSD calculation. The end result is a minimzation of the RMSD calculation by optimizing the rotation of the coordinate set in space to allow for spatial alignment of the coordinate set

RMSD is used to measure structural distance between coordinate sets, i.e. measuring the distance between atoms in a 3D space

HINT: Bio3D functions have various default options that will be used if the
option is not explicitly specified by the user, see help(rmsd) for an example and note that the
input options with an equals sign (e.g. fit=FALSE) have default values.

# Section 5

In this section we perform PCA on the complete collection of Adenylate kinase structures in the
protein data-bank (PDB). Starting from only one PDB identifier (PDB ID 1AKE) we show how to
search the PDB for related structures using BLAST, fetch and align the structures, and finally
calculate the normal modes of each individual structure in order to probe for potential
differences in structural flexibility.

###5.1 Search and retrieve Adenylate kinase structures

```{r}
aa <- get.seq("1ake_A")

# Blast or hmmer search
b <- blast.pdb(aa)

```

Function plot.blast() facilitates the visualization and filtering of the Blast results. It will
attempt to set a seed position to the point of largest drop-off in normalized scores (i.e. the
biggest jump in E-values). In this particular case we specify a cutoff (after initial plotting) of 225
to include only the relevant E.coli structures:

```{r}
# Plot a summary of search results
hits <- plot(b)

head(hits$pdb.id)

```

The Blast search and subsequent filtering identified a total of 39 related PDB structures to our
query sequence. The PDB identifiers of this collection are accessible through the pdb.id
attribute to the hits object (hits$pdb.id). Note that adjusting the cutoff argument (to
plot.blast()) will result in a decrease or increase of hits.

We can now use function get.pdb() and pdbslit() to fetch and parse the identified
structures. Finally, we use pdbaln() to align the PDB structures.

```{r}

# fetch pdbs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip =
TRUE)

```

```{r}

#align structures
pdbs <- pdbaln(files)

#assign names 
ids <- basename.pdb(pdbs$id)

#final plot comparing pdb sequences
plot(pdbs, labels=ids)
```


Schematic representation of alignment. Grey regions depict aligned residues, while
white depict gap regions. The red bar at the top depict sequence conservation. 

### Section 5.4 PCA analysis

Function pca() provides principal component analysis (PCA) of the structure data. PCA is astatistical approach used to transform a data set down to a few important components thatdescribe the directions where there is most variance. In terms of protein structures PCA is used to capture major structural variations within an ensemble of structures.

Function pca() provides principal component analysis (PCA) of the structure data. PCA is a statistical approach used to transform a data set down to a few important components that describe the directions where there is most variance. In terms of protein structures PCA is used to capture major structural variations within an ensemble of structures.

PCA can be performed on the structural ensemble (stored in the pdbs object) with function
pca.xyz(). To obtain meaningful results we first superimpose all structures on the invariant
core (function core.find()).

```{r}

#find most structurally rigid (i.e. the core) sequence amongst all inputs
core <- core.find(pdbs)
```


```{r}
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)

plot(pc.xray)
```

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```












