---
title: "map_colors_function"
author: "Amay"
date: "April 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Fixing function

Ta-da

```{r}

# Old function
map.colors <- function (value,high.low,palette) {
  proportion <- ((value-high.low[1])/(high.low[2]-high.low[1]))
  index <- round ((length(palette)-1)*proportion)+1
  return (palette[index])
}
```


Understanding the function

Every function has 3 things: A name, the input/arguments, and the body. Most have a fourth thing, a return. { } contain the entirety of the function's body, the name is above the body and the inputs are placed inside ( ), also above the body

```{r}

# assume value is a numeric value, replace with x for convenience. Highlighting one instance of word highlights wherever word appears
# high.low appears to take vector inputs based on [] usage, seems to require a two unit vector signifying a range of data

# Description: Map the values of input vector 'x' to a corresponding color from palette
map.colors2 <- function (x,high.low, palette) 

{
  # changed proportion to percent to better reflect purpose of variable (determining where in 'high.low' range the values   of x lie)
  
  percent <- ( (x - high.low[1]) / (high.low[2] - high.low[1]) )
  
  # Find corresponding index position in the color 'palette'
  # Make sure 0 does not become an input, will give error for 'palette' return
  index <- round ( (length(palette) - 1) * percent) + 1
  
  return (palette[index])
}

```


3rd function, final product
```{r}

# Description: Map the values of input vector 'x' to a corresponding color from palette
# added default values for 'high.low' and 'palette'

map.colors3 <- function (x, high.low = range(x), palette = cm.colors(100)) 

{
  # changed proportion to percent to better reflect purpose of variable (determining where in 'high.low' range the values   of x lie)
  
  percent <- ( (x - high.low[1]) / (high.low[2] - high.low[1]) )
  
  # Find corresponding index position in the color 'palette'
  # Make sure 0 does not become an input, will give error for 'palette' return
  
  index <- round ( (length(palette) - 1) * percent) + 1
  
  return (palette[index])
}

```


# Class 6 

## Working with our own functions

My first function

```{r}
add <- function (x, y = 1) {
  # sum input x and y 
  x + y
}
```


Testing first function 
```{r}
add(6)
add(6,8)
add(1:10)
add(2:3,4)
```
```{r}
add( c(1:5), 20 )

```

```{r}
#add(1, y = 'b')
```


```{r}
#add(1, 2, 2)
```

```{r}
## You need a “name”, “arguments” and “body”…
rescale <- function(x) {
rng <-range(x)
(x - rng[1]) / (rng[2] - rng[1])
}
# Test on a small example where you know the answer

```

```{r}
rescale(1:10)
```

Testing a function involves testing it against an ouput you know beforehand, and also what makes the function **fail** allowing you to fix anticipated errors

```{r}
# Test on a small example where you know the answer
rescale(1:10)
# How would you get your function to work here…
#rescale( c(1,2,NA,3,10) )
# What should your function do here?
#rescale( c(1,10,”string”) )
```


Dissect the problem, see where the error starts if you have a vector with NA as a value (in this example)
Analysing it reveals the problem is in the range function, see help section on it


```{r}
rescale2 <- function(x) {
  # changed na.rm default to TRUE, keeping NAs in 
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

Making function fail allows you to set limits on whats permissible for your function

##Function 3

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {

  # if else statement is an example here only, completely not needed
  if(na.rm) {

    rng <-range(x, na.rm=na.rm)
  } 
  else {

    rng <-range(x)
  }

  print("Hello")

  answer <- (x - rng[1]) / (rng[2] - rng[1])

  print("is it me you are looking for?")

  if(plot) {

    plot(answer, typ="b", lwd=4)
  }

  print("I can see it in ...")
  
  return(answer)
  
  #return marks end of function regardless, R will stop executing function after it sees a return. Thus print statement below a return doesn't get read.
}
```

```{r}
rescale3(c(1,2,NA,3,10), plot = T)
``` 


## Section 2B from Class 6
```{r}
library(bio3d)

# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```

```{r}
s1 <- read.pdb("4AKE")
s1

#class of s1 is 'pdb', and the object and functions are coded in such a way that there is a unique 'print.pdb' function associated for objects of class pdb - thus a convenient shortened format of data will be printed as opposed to the entire dataset. To print entire dataset, use print.default('pdb file name')

```

```{r}
library(bio3d)

# Can you improve this analysis code? should be able to take any of the 3 IDs in original code, and make a plot to compare Bfactor
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

# simplify to 1 command for read, one for trim, one for plot
# can't reduce calculations
# turn into a function with name, input, body

```

















